Pulmonary Exacerbations in CF

### 2024-10-22 Scott Sagel/PEx in CF/analysis_for_yaron.qmd
minor updates for Scott

---
title: "Pulmonary Exacerbations in CF"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
    theme:
      light: flatly
      dark: darkly
bibliography: /home/tim/Documents/Miscellaneous/zotero.bib
csl: /home/tim/GitHub/styles/american-medical-association.csl
editor: source
---

```{r setup}
#| include: false
library(dplyr)
library(tidyr)
library(lubridate)
library(gtsummary)
library(ggpubr)
library(rspiro)
knitr::opts_knit$set(root.dir = "/home/tim/OneDrive/Vigers/CF/Scott Sagel/PEx in CF")
# Colors for plots
yellow <- "#f8b80a"
red <- "#e30022"
purple <- "#7b2670"
blue <- "#022451"
```

@@

# Labels
labels <- list(
  "Age group at admission" = "age_group",
  "Age at admission" = "age_at_admit", "Sex" = "sex",
  "Race" = "race", "Ethnicity" = "ethnicity",
  "Azithromycin" = "azithro", "Inhaled antibiotics" = "inhaled_abx",
  "Systemic steroids" = "steroids_system",
  "Source of respiratory culture at admit" = "admit_cx_source",
  "Aspergillus fumigatus" = "fungi___1",
  "Aspergillus (other species)" = "fungi___2",
  "Scedosporium apiospermum" = "fungi___3",
  "Candidia (any species)" = "fungi___4",
  "Other fungus" = "fungi___5",
  "NTM testing performed?" = "ntm_test",
  "Fungal testing performed?" = "fungal_test",
  "Viral testing performed?" = "viral_test",
  "Viral testing results" = "viral_test_results",
  "Influenza A or B" = "viruses___1",
  "Respiratory Syncytial Virus" = "viruses___2",
  "Rhinovirus/enterovirus" = "viruses___3",
  "SARS CoV-2" = "viruses___4",
  "Other virus" = "viruses___5",
  "M. abscessus species" = "ntm___1",
  "M. avium complex species" = "ntm___2",
  "Other NTM species" = "ntm___3",
  "A. xylosoxidans" = "admit_cx_results___1",
  "B. cepacia complex" = "admit_cx_results___2",
  "H. influenzae" = "admit_cx_results___3",
  "MRSA" = "admit_cx_results___4",
  "Methicillin-susceptible Staph aureus" = "admit_cx_results___5",
  "P. aeruginosa" = "admit_cx_results___6",
  "P. aeruginosa mucoid" = "admit_cx_results___7",
  "S. maltophilia" = "admit_cx_results___8",
  "Other bacteria" = "admit_cx_results___9",
  "On CFTR Modulator?" = "cftr_mod",
  "CFTR Modulator Type" = "cftr_mod_spec",
  "Cultures Tested" = "redcap_data_access_group",
  "Length of Admission" = "days_admit",
  "Systemic Steroids" = "steroids_system",
  "Oral Steroids" = "steroids_oral",
  "IV Steroids" = "steroids_iv",
  "Baseline FEV1pp (2022)" = "ppfev1_baseline",
  "Admission FEV1pp (2022)" = "ppfev1_admit",
  "Discharge FEV1pp (2022)" = "ppfev1_discharge",
  "Change in mean FEV1pp from admit to discharge" = "ppfev1_change"

@@

labels <- as.list(setNames(names(labels), labels))
# Format
  select(
    sid,
    redcap_data_access_group, year, sex, age_group, age_at_admit, race,
    ethnicity, Genotype, Insurance, cftr_mod, cftr_mod_spec,
    Trikafta, azithro, inhaled_abx, steroids_system, admit_cx_source,
    ppfev1_baseline, ppfev1_admit, ppfev1_discharge, days_admit,
    fungal_test, fungi___1:fungi___5,
    ntm_test, ntm___1:ntm___3,
    viral_test, viruses___1:viruses___5,
    admit_cx_results___1:admit_cx_results___9,
    contains("steroids")
  filter(!is.na(year))
# Write for Scott and Yaron to check
write.csv(df,file = "./Data_Clean/analysis_dataset.csv",row.names = F,na = "")
```

@@

```{r}
#| label: tbl-table-1
#| tbl-cap: Participant characteristics at admission
df %>%
  select(redcap_data_access_group:days_admit) %>%
  tbl_summary(
    by = year, missing_text = "(Missing)", label = labels,
    statistic =
      list(
        all_continuous() ~ "{median} [{p25}, {p75}]",
        c(age_at_admit, ppfev1_baseline, ppfev1_admit, ppfev1_discharge) ~
          "{mean} ({sd})"
      )
  ) %>%
  add_p(
    test =
      list(
        c(age_at_admit, ppfev1_baseline, ppfev1_admit, ppfev1_discharge) ~
          "t.test"
      ),
    test.args = all_tests("fisher.test") ~ list(simulate.p.value = T)
  ) %>%
  separate_p_footnotes()
```

@@

```{r}
#| label: tbl-fungi
#| tbl-cap: Fungi
df %>%
  filter(fungal_test == "Yes") %>%
  select(year, redcap_data_access_group, fungi___1:fungi___5) %>%
  tbl_summary(by = year, missing_text = "Missing", label = labels) %>%
  add_overall() %>%
  add_p() %>%
  separate_p_footnotes()
```

```{r}
#| label: tbl-fungi-center
#| tbl-cap: Fungi by center
#| message: false
df %>%
  filter(fungal_test == "Yes") %>%
  select(year, redcap_data_access_group, fungi___1:fungi___5) %>%
  tbl_strata(
    strata = redcap_data_access_group,
    .tbl_fun =
      ~ .x %>%
        tbl_summary(by = year, missing_text = "Missing", label = labels) %>%
        add_overall() %>%
        add_p() %>%
        separate_p_footnotes()
  )
```

```{r}
#| label: tbl-fungi-cftr-mod
#| tbl-cap: Fungi by modulator status
#| message: false
df %>%
  filter(fungal_test == "Yes") %>%
  select(year, cftr_mod, fungi___1:fungi___5) %>%
  tbl_strata(
    strata = year,
    .tbl_fun =
      ~ .x %>%
        tbl_summary(by = cftr_mod, missing_text = "Missing", label = labels) %>%
        add_overall() %>%
        add_p() %>%
        separate_p_footnotes()
  )
```

```{r}
#| label: tbl-fungi-cftr-mod-spec
#| tbl-cap: Fungi by specific modulator status
#| message: false
df %>%
  filter(fungal_test == "Yes") %>%
  select(year, cftr_mod_spec, fungi___1:fungi___5) %>%
  tbl_strata(
    strata = year,
    .tbl_fun =
      ~ .x %>%
        tbl_summary(
          by = cftr_mod_spec, missing_text = "Missing",
          label = labels
        ) %>%
        add_overall() %>%
        add_p() %>%
        separate_p_footnotes()
  )
```

```{r}
#| label: tbl-fungi-trikafta
#| tbl-cap: Fungi by Trikafta status
#| message: false
df %>%
  filter(fungal_test == "Yes") %>%
  select(year, Trikafta, fungi___1:fungi___5) %>%
  tbl_strata(
    strata = year,
    .tbl_fun =
      ~ .x %>%
        tbl_summary(by = Trikafta, missing_text = "Missing", label = labels) %>%
        add_overall() %>%
        add_p() %>%
        separate_p_footnotes()
  )
```

@@

```{r}
#| label: tbl-ntm
#| tbl-cap: NTM
df %>%
  filter(ntm_test == "Yes") %>%
  select(year, redcap_data_access_group, ntm___1:ntm___3) %>%
  tbl_summary(by = year, missing_text = "Missing", label = labels) %>%
  add_overall() %>%
  add_p() %>%
  separate_p_footnotes()
```

```{r}
#| label: tbl-ntm-center
#| tbl-cap: NTM by center
#| message: false
df %>%
  filter(ntm_test == "Yes") %>%
  select(year, redcap_data_access_group, ntm___1:ntm___3) %>%
  tbl_strata(
    strata = redcap_data_access_group,
    .tbl_fun =
      ~ .x %>%
        tbl_summary(by = year, missing_text = "Missing", label = labels) %>%
        add_overall() %>%
        add_p() %>%
        separate_p_footnotes()
  )
```

```{r}
#| label: tbl-ntm-cftr-mod
#| tbl-cap: NTM by modulator status
#| message: false
df %>%
  filter(ntm_test == "Yes") %>%
  select(year, cftr_mod, ntm___1:ntm___3) %>%
  tbl_strata(
    strata = year,
    .tbl_fun =
      ~ .x %>%
        tbl_summary(
          by = cftr_mod, missing_text = "Missing",
          label = labels
        ) %>%
        add_overall() %>%
        add_p() %>%
        separate_p_footnotes()
  )
```

```{r}
#| label: tbl-ntm-cftr-mod-spec
#| tbl-cap: NTM by specific modulator status
#| message: false
df %>%
  filter(ntm_test == "Yes") %>%
  select(year, cftr_mod_spec, ntm___1:ntm___3) %>%
  tbl_strata(
    strata = year,
    .tbl_fun =
      ~ .x %>%
        tbl_summary(
          by = cftr_mod_spec, missing_text = "Missing",
          label = labels
        ) %>%
        add_overall() %>%
        add_p() %>%
        separate_p_footnotes()
  )
```

```{r}
#| label: tbl-ntm-trikafta
#| tbl-cap: NTM by Trikafta status
#| message: false
df %>%
  filter(ntm_test == "Yes") %>%
  select(year, Trikafta, ntm___1:ntm___3) %>%
  tbl_strata(
    strata = year,
    .tbl_fun =
      ~ .x %>%
        tbl_summary(by = Trikafta, missing_text = "Missing", label = labels) %>%
        add_overall() %>%
        add_p() %>%
        separate_p_footnotes()
  )
```

@@

```{r}
#| label: tbl-viruses
#| tbl-cap: Viruses
df %>%
  filter(viral_test == "Yes") %>%
  select(year, redcap_data_access_group, viruses___1:viruses___5) %>%
  tbl_summary(by = year, missing_text = "Missing", label = labels) %>%
  add_overall() %>%
  add_p() %>%
  separate_p_footnotes()
```

```{r}
#| label: tbl-viruses-cftr-mod
#| tbl-cap: Viruses by modulator status
#| message: false
df %>%
  filter(viral_test == "Yes") %>%
  select(year, cftr_mod, viruses___1:viruses___5) %>%
  tbl_strata(
    strata = year,
    .tbl_fun =
      ~ .x %>%
        tbl_summary(by = cftr_mod, missing_text = "Missing", label = labels) %>%
        add_overall() %>%
        add_p() %>%
        separate_p_footnotes()
  )
```

```{r}
#| label: tbl-viruses-cftr-mod-spec
#| tbl-cap: Viruses by specific modulator status
#| message: false
df %>%
  filter(viral_test == "Yes") %>%
  select(year, cftr_mod_spec, viruses___1:viruses___5) %>%
  tbl_strata(
    strata = year,
    .tbl_fun =
      ~ .x %>%
        tbl_summary(
          by = cftr_mod_spec, missing_text = "Missing",
          label = labels
        ) %>%
        add_overall() %>%
        add_p() %>%
        separate_p_footnotes()
  )
```

```{r}
#| label: tbl-viruses-trikafta
#| tbl-cap: Viruses by Trikafta status
#| message: false
df %>%
  filter(viral_test == "Yes") %>%
  select(year, Trikafta, viruses___1:viruses___5) %>%
  tbl_strata(
    strata = year,
    .tbl_fun =
      ~ .x %>%
        tbl_summary(by = Trikafta, missing_text = "Missing", label = labels) %>%
        add_overall() %>%
        add_p() %>%
        separate_p_footnotes()
  )
```

@@

```{r}
#| label: tbl-bacteria
#| tbl-cap: Bacteria
df %>%
  select(
    year, redcap_data_access_group,
    admit_cx_results___1:admit_cx_results___9
  ) %>%
  tbl_summary(by = year, missing_text = "Missing", label = labels) %>%
  add_overall() %>%
  add_p() %>%
  separate_p_footnotes()
```

```{r}
#| label: tbl-bacteria-center
#| tbl-cap: Bacteria by center
#| message: false
df %>%
  select(
    year, redcap_data_access_group, admit_cx_results___1:admit_cx_results___9
  ) %>%
  tbl_strata(
    strata = redcap_data_access_group,
    .tbl_fun =
      ~ .x %>%
        tbl_summary(by = year, missing_text = "Missing", label = labels) %>%
        add_overall() %>%
        add_p() %>%
        separate_p_footnotes()
  )
```

```{r}
#| label: tbl-bacteria-cftr-mod
#| tbl-cap: Bacteria by modulator status
#| message: false
df %>%
  select(year, cftr_mod, admit_cx_results___1:admit_cx_results___9) %>%
  tbl_strata(
    strata = year,
    .tbl_fun =
      ~ .x %>%
        tbl_summary(
          by = cftr_mod, missing_text = "Missing",
          label = labels
        ) %>%
        add_overall() %>%
        add_p() %>%
        separate_p_footnotes()
  )
```

```{r}
#| label: tbl-bacteria-cftr-mod-spec
#| tbl-cap: Bacteria by specfic modulator status
#| message: false
df %>%
  select(year, cftr_mod_spec, admit_cx_results___1:admit_cx_results___9) %>%
  tbl_strata(
    strata = year,
    .tbl_fun =
      ~ .x %>%
        tbl_summary(
          by = cftr_mod_spec, missing_text = "Missing",
          label = labels
        ) %>%
        add_overall() %>%
        add_p() %>%
        separate_p_footnotes()
  )
```

```{r}
#| label: tbl-bacteria-trikafta
#| tbl-cap: Bacteria by Trikafta status
#| message: false
df %>%
  select(year, Trikafta, admit_cx_results___1:admit_cx_results___9) %>%
  tbl_strata(
    strata = year,
    .tbl_fun =
      ~ .x %>%
        tbl_summary(by = Trikafta, missing_text = "Missing", label = labels) %>%
        add_overall() %>%
        add_p() %>%
        separate_p_footnotes()
  )
```

@@

```{r}
#| label: fig-num-hosps
#| fig-cap: Number of hospitalizations
df %>%
  group_by(year) %>%
  summarise(n = n()) %>%
  ggplot(aes(x = factor(year), y = n, fill = factor(year))) +
  geom_bar(position = "dodge", stat = "identity") +
  geom_label(aes(label = n), nudge_y = -20, fill = "white") +
  ylab("Num. of Hospitalizations") +
  theme_bw() +
  theme(legend.position = "none", axis.title.x = element_blank()) +
  scale_fill_manual(values = c(purple, yellow))
```

```{r}
#| label: fig-perc-hosp-2-plus
#| fig-cap: Percent hospitalized 2 or more times in both years
df %>%
  group_by(sid, year) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(year) %>%
  summarise(
    num = sum(n >= 2), total = n(),
    perc = mean(n >= 2) * 100, .groups = "drop"
  ) %>%
  ggplot(aes(x = factor(year), y = perc, fill = factor(year))) +
  geom_bar(position = "dodge", stat = "identity") +
  geom_label(aes(label = scales::percent(perc, scale = 1, accuracy = 0.1)),
    nudge_y = -5, fill = "white"
  ) +
  geom_label(aes(label = paste0(num, "/", total)),
    nudge_y = -8, fill = "white"
  ) +
  geom_signif(
    xmin = 1, xmax = 2, y_position = 50, map_signif_level = T,
    annotation = "<0.001"
  ) +
  xlab("Site") +
  ylab("Percent With >= 2 Hospitalizations") +
  theme_bw() +
  theme(legend.position = "none") +
  scale_fill_manual(values = c(purple, yellow))
```

@@

Hospitalizations within a person were considered independent in the table below.

>>>>

---
title: "Pulmonary Exacerbations in CF"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
    theme:
      light: flatly
      dark: darkly
editor: source
---

```{r setup}
#| include: false
library(dplyr)
library(tidyr)
library(lubridate)
library(gtsummary)
library(ggpubr)
library(rspiro)
library(patchwork)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Scott Sagel/PEx in CF",
  "Windows" = "C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Scott Sagel/PEx in CF",
  "Linux" = "/home/timvigers/OneDrive/Vigers/CF/Scott Sagel/PEx in CF"
)
github_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/GitHub",
  "Windows" = "C:/Users/timvigers/GitHub",
  "Linux" = "/home/timvigers/GitHub"
)
knitr::opts_knit$set(root.dir = home_dir)
# Colors for plots
yellow <- "#f8b80a"
red <- "#e30022"
purple <- "#7b2670"
blue <- "#022451"
```

@@

df$Insurance <- factor(df$Insurance,
    "Medicaid", "Medicaid/None", "Medicaid/Other", "Medicare", "Medicare/Other",
    "None", "Other", "Private insurance", "Private insurance/Medicaid",
    "Private insurance/Other",
    "Private insurance/Tricare/other military health plan",
    "Tricare/other military health plan"
    "Medicaid", "Medicaid", "Medicaid", "Medicare", "Medicare", "None/Other",
    "None/Other", "Private insurance", "Private insurance", "Private insurance",
    "Tricare/other military health plan", "Tricare/other military health plan"

@@

df$admit_cx_source <- factor(df$admit_cx_source,
  levels = 1:4,
    "Bronchoscopy", "Expectorated sputum", "Induced sputum", "Throat swab"
# Combine species
df$any_fungi <- apply(df, 1, function(r) {
  r <- as.numeric(r[paste0("fungi___", 1:5)])
  any(r == 1)
})
df$any_fungi <- factor(df$any_fungi, levels = c(F, T), labels = c("No", "Yes"))
df$any_ntm <- apply(df, 1, function(r) {
  r <- as.numeric(r[paste0("ntm___", 1:3)])
  any(r == 1)
})
df$any_ntm <- factor(df$any_ntm, levels = c(F, T), labels = c("No", "Yes"))
df$any_virus <- apply(df, 1, function(r) {
  r <- as.numeric(r[paste0("viruses___", 1:5)])
  any(r == 1)
})
df$any_virus <- factor(df$any_virus, levels = c(F, T), labels = c("No", "Yes"))
df$any_bact <- apply(df, 1, function(r) {
  r <- as.numeric(r[paste0("admit_cx_results___", 1:9)])
  any(r == 1)
})
df$any_bact <- factor(df$any_bact, levels = c(F, T), labels = c("No", "Yes"))

df$no_bact <- apply(df, 1, function(r) {
  r <- as.numeric(r[paste0("admit_cx_results___", 1:8)])
  all(r == 0)
})
df$no_bact <- factor(df$no_bact, levels = c(F, T), labels = c("No", "Yes"))
# Change in FEV1 from admission to discharge
df$ppfev1_change <- df$ppfev1_discharge - df$ppfev1_admit
# Format fungi, NTM, etc. columns
df <- df %>%
  mutate(
    across(
      c(
        fungi___1:fungi___5, ntm___1:ntm___3, viruses___1:viruses___5,
        admit_cx_results___1:admit_cx_results___9
      ),
      ~ factor(as.numeric(.x),
        levels = 0:1,
        labels = c("Not detected", "Detected")
      )
    )
  )
# Labels
labels <- list(
  "Age group at admission" = "age_group",
  "Age at admission" = "age_at_admit", "Sex" = "sex",
  "Race" = "race", "Ethnicity" = "ethnicity",
  "Azithromycin" = "azithro", "Inhaled antibiotics" = "inhaled_abx",
  "Systemic steroids" = "steroids_system",
  "Source of respiratory culture at admit" = "admit_cx_source",
  "Aspergillus fumigatus" = "fungi___1",
  "Aspergillus (other species)" = "fungi___2",
  "Scedosporium apiospermum" = "fungi___3",
  "Candidia (any species)" = "fungi___4",
  "Other fungus" = "fungi___5",
  "Any fungi?" = "any_fungi",
  "NTM testing performed?" = "ntm_test",
  "Fungal testing performed?" = "fungal_test",
  "Viral testing performed?" = "viral_test",
  "Viral testing results" = "viral_test_results",
  "Influenza A or B" = "viruses___1",
  "Respiratory Syncytial Virus" = "viruses___2",
  "Rhinovirus/enterovirus" = "viruses___3",
  "SARS CoV-2" = "viruses___4",
  "Other virus" = "viruses___5",
  "Any virus?" = "any_virus",
  "M. abscessus species" = "ntm___1",
  "M. avium complex species" = "ntm___2",
  "Other NTM species" = "ntm___3",
  "Any NTM?" = "any_ntm",
  "A. xylosoxidans" = "admit_cx_results___1",
  "B. cepacia complex" = "admit_cx_results___2",
  "H. influenzae" = "admit_cx_results___3",
  "MRSA" = "admit_cx_results___4",
  "Methicillin-susceptible Staph aureus" = "admit_cx_results___5",
  "P. aeruginosa" = "admit_cx_results___6",
  "P. aeruginosa mucoid" = "admit_cx_results___7",
  "S. maltophilia" = "admit_cx_results___8",
  "Other bacteria" = "admit_cx_results___9",
  "Any bacteria?" = "any_bact",
  "No CF bacteria" = "no_bact",
  "On CFTR Modulator?" = "cftr_mod",
  "CFTR Modulator Type" = "cftr_mod_spec",
  "Site" = "redcap_data_access_group",
  "Length of Admission" = "days_admit",
  "Systemic Steroids" = "steroids_system",
  "Oral Steroids" = "steroids_oral",
  "IV Steroids" = "steroids_iv",
  "Baseline FEV1pp" = "ppfev1_baseline",
  "Admission FEV1pp" = "ppfev1_admit",
  "Discharge FEV1pp" = "ppfev1_discharge",
  "Change in mean FEV1pp from admit to discharge" = "ppfev1_change"
)
labels <- as.list(setNames(names(labels), labels))
# Format
df <- df %>%
  select(
    sid,
    redcap_data_access_group, year, sex, age_group, age_at_admit, race,
    ethnicity, Genotype, Insurance, cftr_mod, cftr_mod_spec,
    Trikafta, azithro, inhaled_abx, steroids_system, admit_cx_source,
    ppfev1_baseline, ppfev1_admit, ppfev1_discharge, ppfev1_change,
    days_admit, fungal_test, fungi___1:fungi___5, any_fungi,
    ntm_test, ntm___1:ntm___3, any_ntm,
    viral_test, viruses___1:viruses___5, any_virus,
    admit_cx_results___1:admit_cx_results___9, any_bact, no_bact,
    contains("steroids")
  ) %>%
  filter(!is.na(year))
```

@@

```{r}
#| label: tbl-table-1
#| tbl-cap: Participant characteristics overall
df %>%
  group_by(sid) %>%
  filter(row_number() == 1) %>%
  ungroup() %>%
  select(redcap_data_access_group, sex, race, ethnicity, Genotype) %>%
  tbl_summary(missing_text = "(Missing)", label = labels)
```

```{r}
#| label: tbl-table-1-year
#| tbl-cap: Participant characteristics by year
df %>%
  group_by(sid, year) %>%
  filter(row_number() == 1) %>%
  ungroup() %>%
  select(year, redcap_data_access_group, sex, race, ethnicity, Genotype) %>%
  tbl_summary(by = year, missing_text = "(Missing)", label = labels)
```

```{r}
#| label: tbl-table-1-indep
#| tbl-cap: Participant characteristics at admission by year
df %>%
  select(redcap_data_access_group:days_admit) %>%
  tbl_summary(
    by = year, missing_text = "(Missing)", label = labels,
    statistic =
      list(
        all_continuous() ~ "{median} [{p25}, {p75}]",
        c(
          age_at_admit, ppfev1_baseline, ppfev1_admit, ppfev1_discharge,
          ppfev1_change
        ) ~
          "{mean} ({sd})"
      )
  ) %>%
  add_p(
    test =
      list(
        c(age_at_admit, ppfev1_baseline, ppfev1_admit, ppfev1_discharge) ~
          "t.test"
      ),
    test.args = all_tests("fisher.test") ~ list(simulate.p.value = T)
  ) %>%
  separate_p_footnotes()
```

@@

```{r}
#| label: tbl-fungi
#| tbl-cap: Fungi
df %>%
  filter(fungal_test == "Yes") %>%
  select(year, redcap_data_access_group, fungi___1:fungi___5, any_fungi) %>%
  tbl_summary(by = year, missing_text = "Missing", label = labels) %>%
  add_overall() %>%
  add_p() %>%
  separate_p_footnotes()
```

```{r}
#| label: tbl-fungi-center
#| tbl-cap: Fungi by center
#| message: false
df %>%
  filter(fungal_test == "Yes") %>%
  select(year, redcap_data_access_group, fungi___1:fungi___5, any_fungi) %>%
  tbl_strata(
    strata = redcap_data_access_group,
    .tbl_fun =
      ~ .x %>%
        tbl_summary(by = year, missing_text = "Missing", label = labels) %>%
        add_overall() %>%
        add_p() %>%
        separate_p_footnotes()
  )
```

```{r}
#| label: tbl-fungi-cftr-mod
#| tbl-cap: Fungi by modulator status
#| message: false
df %>%
  filter(fungal_test == "Yes") %>%
  select(year, cftr_mod, fungi___1:fungi___5, any_fungi) %>%
  tbl_strata(
    strata = year,
    .tbl_fun =
      ~ .x %>%
        tbl_summary(by = cftr_mod, missing_text = "Missing", label = labels) %>%
        add_overall() %>%
        add_p() %>%
        separate_p_footnotes()
  )
```

```{r}
#| label: tbl-fungi-cftr-mod-spec
#| tbl-cap: Fungi by specific modulator status
#| message: false
df %>%
  filter(fungal_test == "Yes") %>%
  select(year, cftr_mod_spec, fungi___1:fungi___5, any_fungi) %>%
  tbl_strata(
    strata = year,
    .tbl_fun =
      ~ .x %>%
        tbl_summary(
          by = cftr_mod_spec, missing_text = "Missing",
          label = labels
        ) %>%
        add_overall() %>%
        add_p() %>%
        separate_p_footnotes()
  )
```

```{r}
#| label: tbl-fungi-trikafta
#| tbl-cap: Fungi by Trikafta status
#| message: false
df %>%
  filter(fungal_test == "Yes") %>%
  select(year, Trikafta, fungi___1:fungi___5, any_fungi) %>%
  tbl_strata(
    strata = year,
    .tbl_fun =
      ~ .x %>%
        tbl_summary(by = Trikafta, missing_text = "Missing", label = labels) %>%
        add_overall() %>%
        add_p() %>%
        separate_p_footnotes()
  )
```

@@

```{r}
#| label: tbl-ntm
#| tbl-cap: NTM
df %>%
  filter(ntm_test == "Yes") %>%
  select(year, redcap_data_access_group, ntm___1:ntm___3, any_ntm) %>%
  tbl_summary(by = year, missing_text = "Missing", label = labels) %>%
  add_overall() %>%
  add_p() %>%
  separate_p_footnotes()
```

```{r}
#| label: tbl-ntm-center
#| tbl-cap: NTM by center
#| message: false
df %>%
  filter(ntm_test == "Yes") %>%
  select(year, redcap_data_access_group, ntm___1:ntm___3, any_ntm) %>%
  tbl_strata(
    strata = redcap_data_access_group,
    .tbl_fun =
      ~ .x %>%
        tbl_summary(by = year, missing_text = "Missing", label = labels) %>%
        add_overall() %>%
        add_p() %>%
        separate_p_footnotes()
  )
```

```{r}
#| label: tbl-ntm-cftr-mod
#| tbl-cap: NTM by modulator status
#| message: false
df %>%
  filter(ntm_test == "Yes") %>%
  select(year, cftr_mod, ntm___1:ntm___3, any_ntm) %>%
  tbl_strata(
    strata = year,
    .tbl_fun =
      ~ .x %>%
        tbl_summary(
          by = cftr_mod, missing_text = "Missing",
          label = labels
        ) %>%
        add_overall() %>%
        add_p() %>%
        separate_p_footnotes()
  )
```

```{r}
#| label: tbl-ntm-cftr-mod-spec
#| tbl-cap: NTM by specific modulator status
#| message: false
df %>%
  filter(ntm_test == "Yes") %>%
  select(year, cftr_mod_spec, ntm___1:ntm___3, any_ntm) %>%
  tbl_strata(
    strata = year,
    .tbl_fun =
      ~ .x %>%
        tbl_summary(
          by = cftr_mod_spec, missing_text = "Missing",
          label = labels
        ) %>%
        add_overall() %>%
        add_p() %>%
        separate_p_footnotes()
  )
```

```{r}
#| label: tbl-ntm-trikafta
#| tbl-cap: NTM by Trikafta status
#| message: false
df %>%
  filter(ntm_test == "Yes") %>%
  select(year, Trikafta, ntm___1:ntm___3, any_ntm) %>%
  tbl_strata(
    strata = year,
    .tbl_fun =
      ~ .x %>%
        tbl_summary(by = Trikafta, missing_text = "Missing", label = labels) %>%
        add_overall() %>%
        add_p() %>%
        separate_p_footnotes()
  )
```

@@

```{r}
#| label: tbl-viruses
#| tbl-cap: Viruses
df %>%
  filter(viral_test == "Yes") %>%
  select(year, redcap_data_access_group, viruses___1:viruses___5, any_virus) %>%
  tbl_summary(by = year, missing_text = "Missing", label = labels) %>%
  add_overall() %>%
  add_p() %>%
  separate_p_footnotes()
```

```{r}
#| label: tbl-viruses-cftr-mod
#| tbl-cap: Viruses by modulator status
#| message: false
df %>%
  filter(viral_test == "Yes") %>%
  select(year, cftr_mod, viruses___1:viruses___5, any_virus) %>%
  tbl_strata(
    strata = year,
    .tbl_fun =
      ~ .x %>%
        tbl_summary(by = cftr_mod, missing_text = "Missing", label = labels) %>%
        add_overall() %>%
        add_p() %>%
        separate_p_footnotes()
  )
```

```{r}
#| label: tbl-viruses-cftr-mod-spec
#| tbl-cap: Viruses by specific modulator status
#| message: false
df %>%
  filter(viral_test == "Yes") %>%
  select(year, cftr_mod_spec, viruses___1:viruses___5, any_virus) %>%
  tbl_strata(
    strata = year,
    .tbl_fun =
      ~ .x %>%
        tbl_summary(
          by = cftr_mod_spec, missing_text = "Missing",
          label = labels
        ) %>%
        add_overall() %>%
        add_p() %>%
        separate_p_footnotes()
  )
```

```{r}
#| label: tbl-viruses-trikafta
#| tbl-cap: Viruses by Trikafta status
#| message: false
df %>%
  filter(viral_test == "Yes") %>%
  select(year, Trikafta, viruses___1:viruses___5, any_virus) %>%
  tbl_strata(
    strata = year,
    .tbl_fun =
      ~ .x %>%
        tbl_summary(by = Trikafta, missing_text = "Missing", label = labels) %>%
        add_overall() %>%
        add_p() %>%
        separate_p_footnotes()
  )
```

@@

```{r}
#| label: tbl-bacteria
#| tbl-cap: Bacteria
df %>%
  select(
    year, redcap_data_access_group,
    admit_cx_results___1:admit_cx_results___9, any_bact, no_bact
  ) %>%
  tbl_summary(by = year, missing_text = "Missing", label = labels) %>%
  add_overall() %>%
  add_p() %>%
  separate_p_footnotes()
```

```{r}
#| label: tbl-bacteria-center
#| tbl-cap: Bacteria by center
#| message: false
df %>%
  select(
    year, redcap_data_access_group, admit_cx_results___1:admit_cx_results___9,
    any_bact, no_bact
  ) %>%
  tbl_strata(
    strata = redcap_data_access_group,
    .tbl_fun =
      ~ .x %>%
        tbl_summary(by = year, missing_text = "Missing", label = labels) %>%
        add_overall() %>%
        add_p() %>%
        separate_p_footnotes()
  )
```

```{r}
#| label: tbl-bacteria-cftr-mod
#| tbl-cap: Bacteria by modulator status
#| message: false
df %>%
  select(
    year, cftr_mod, admit_cx_results___1:admit_cx_results___9, any_bact, no_bact
  ) %>%
  tbl_strata(
    strata = year,
    .tbl_fun =
      ~ .x %>%
        tbl_summary(
          by = cftr_mod, missing_text = "Missing",
          label = labels
        ) %>%
        add_overall() %>%
        add_p() %>%
        separate_p_footnotes()
  )
```

```{r}
#| label: tbl-bacteria-cftr-mod-spec
#| tbl-cap: Bacteria by specfic modulator status
#| message: false
df %>%
  select(
    year, cftr_mod_spec, admit_cx_results___1:admit_cx_results___9, any_bact,
    no_bact
  ) %>%
  tbl_strata(
    strata = year,
    .tbl_fun =
      ~ .x %>%
        tbl_summary(
          by = cftr_mod_spec, missing_text = "Missing",
          label = labels
        ) %>%
        add_overall() %>%
        add_p() %>%
        separate_p_footnotes()
  )
```

```{r}
#| label: tbl-bacteria-trikafta
#| tbl-cap: Bacteria by Trikafta status
#| message: false
df %>%
  select(
    year, Trikafta, admit_cx_results___1:admit_cx_results___9, any_bact, no_bact
  ) %>%
  tbl_strata(
    strata = year,
    .tbl_fun =
      ~ .x %>%
        tbl_summary(by = Trikafta, missing_text = "Missing", label = labels) %>%
        add_overall() %>%
        add_p() %>%
        separate_p_footnotes()
  )
```

@@

```{r}
#| label: fig-num-hosps
#| fig-cap: Number of hospitalizations
fig1 <- df %>%
  group_by(year) %>%
  summarise(n = n()) %>%
  ggplot(aes(x = factor(year), y = n, fill = factor(year))) +
  geom_bar(position = "dodge", stat = "identity") +
  geom_label(aes(label = n), nudge_y = -20, fill = "white") +
  ylab("Num. of Hospitalizations") +
  theme_bw() +
  theme(legend.position = "none", axis.title.x = element_blank()) +
  scale_fill_manual(values = c(purple, yellow))
fig1
```

```{r}
#| label: fig-perc-hosp-2-plus
#| fig-cap: Percent hospitalized 2 or more times in both years
fig2 <- df %>%
  group_by(sid, year) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(year) %>%
  summarise(
    num = sum(n >= 2), total = n(),
    perc = mean(n >= 2) * 100, .groups = "drop"
  ) %>%
  ggplot(aes(x = factor(year), y = perc, fill = factor(year))) +
  geom_bar(position = "dodge", stat = "identity") +
  geom_label(aes(label = scales::percent(perc, scale = 1, accuracy = 0.1)),
    nudge_y = -5, fill = "white"
  ) +
  geom_label(aes(label = paste0(num, "/", total)),
    nudge_y = -8, fill = "white"
  ) +
  geom_signif(
    xmin = 1, xmax = 2, y_position = 50, map_signif_level = T,
    annotation = "<0.001"
  ) +
  ylab("Percent With >= 2 Hospitalizations") +
  theme_bw() +
  theme(legend.position = "none", axis.title.x = element_blank()) +
  scale_fill_manual(values = c(purple, yellow))
fig2
```

```{r}
#| label: fig-panel
#| fig-cap: Hospitalizations
fig1 + fig2
```

@@

Hospitalizations within a person were considered independent in the tables below.
